Operational Feature Selection with Gaussian Mixture Models for remote sensing images

### 2016-10-13 paper.tex
DOC: tipo

\IEEEPARstart{W}{ith} the increasing number of remote sensing missions, the quantity of available data for a given landscape becomes larger and larger. Satellite missions produce a huge amount of data on a regularly (daily) basis. From 2018, the EnMAP (Environmental Mapping and Analysis Program) satellites missioned by the German space agency will produce images with 244 spectral bands with a resolution of 30x30m with a frequency of revisit of 4 days\cite{Müller09enmap}. The Hyperspectral Infrared Imager (HyspIRI) of NASA will also deliver images with 212 spectral bands. Additionally to hyperspectral data, the amount of available hypertemporal data increases a lot. For instance, the European satellites Sentinel-2 were launched successfully recently and 2 Terabytes of data will be released every days~\cite{drusch2012sentinel}. This high volume of earth observation data provides an accurate information of land state and functioning, and help to improve the understanding of the earth~\cite{rs1010001}. However, processing such data is more and more challenging because of statistical and computational issues.

In the spectral or temporal domain, a pixel is represented by a vector
for   wich  each   component   corresponds   to  a   spectral/temporal
measurement.  The  size of  the vector  is the  number of  spectral or
temporal  measurements.   For  hyperspectral images,  this  number  is
typically   about   several   hundreds  while   for   the   Sentinel-2
multitemporal images, the number of spectro-temporal measurement for a
given  year  is approximately  one  thousand.   When working  in  high
dimensional  space,  statistical  methods  made for  low  or  moderate
dimensional  space do  not  adapt  well.  For  instance,  the rate  of
convergence of the statistical estimation decreases when the dimension
grows while conjointly the number of parameters to estimate increases,
making    the    estimation    of   the    model    parameters    very
difficult~\cite{donoho}.  Consequently,  with a limited  training set,
beyond a certain limit, the classification accuracy actually decreases
as the number of features increases~\cite{hughes}.  For the purpose of
classification,  these  problems are  related  to  the \emph{curse  of
  dimensionality}. This is a major drawback in remote sensing since it
is  difficult to  collect a  large and  an accurate  ground-truth.  An
intensive work has  been performed in the remote  sensing community to
build  accurate classifiers  for  high  dimensional images.   Bayesian
models~\cite{book:landgrebe}, feature extraction and feature reduction
techniques~\cite{book:landgrebe,DR:guided:tour},                random
forest~\cite{1396322},   neural  networks~\cite{5411821}   and  kernel
methods~\cite{kernel:methods:rs}  have   been  investigated   for  the
classification of such images.

The  size of  the data  is increasing  with respect  to the  number of
measurement per  pixel. The data  volume of an hyperspectral  image is
typically    several    Gigabytes     per    acquisition    ($\approx$
300km$^2$). Multitemporal data are  now available freely from internet
stream    (see    for    instance     the    Copernicus    data    hub
\url{https://cophub.copernicus.eu/}).  This very  large volume of data
requires   specific   computing  infrastructure.    High   performance
computing   is   actually   investigated   by   the   remote   sensing
community~\cite{christophe2011remote,plaza2011high}.  Main  issues are
related to the use of  parallel approach (multicore, CPU, clusters) to
improve the  processing time, and  to the use of  streaming techniques
when data  does not fit in  memory.  One popular open  source software
solution  is  the Orfeo  Toolbox,  developed  by the  French  National
Agency~\cite{christophe2008orfeo}.   Streaming and  parallel computing
are  convenientely  proposed  to  the  user  through  several  modules
``\emph{ready to use}''.

@@

There are two ways to reduce dimension~\cite[Chapter 1]{Guyon:2006:FEF:1208773}: features extraction and feature selection. Feature extraction means reformulate and summarize the information by creating new features in combining the existing ones, it is sometimes referred to \emph{feature construction}. Linear combination of the initial features can be extracted using Principal Component Analysis (PCA)~\cite{jimenez1998supervised} or Independent Component Analysis \cite{villa2011hyperspectral}. Supervised extraction method has also been investigated such as Fischer discriminant analysis and decision boundary feature extraction~\cite{book:landgrebe}. To the contrary, features selection selects a subset of existing features identified as the most relevant for a given criterion. It has the additional advantage to be much more interpretable for the end-user.

@@

    For mixture models, it is assumed  that a given sample $\mathbf{x}$ is
    the realization  of a  random vector which  distribution is  a mixture
    (convex     combination)     of      several     class     conditioned
    distribution~\cite{Fraley00model-basedclustering}:
    \begin{equation}
        p(\mathbf{x}) = \sum_{c=1}^{C} \pi_c f_c(\mathbf{x}|\theta)
    \end{equation}
    where $\pi_c$ is  the prior, \emph{i.e.}, the  proportion of class
    $c$  and  $f_c$  a  parametric density  function  parametrized  by
    $\theta$.

@@

        The    KL   divergence    is   not    symmetric,   \emph{i.e.}
        $KL_{cc^\prime} \ne KL_{c^\prime c}$.  A symmetrized version  is used to
        compute the criterion function:
        \begin{align}\label{eq:skl}
          SKL_{cc^\prime} &=KL_{cc^\prime} + KL_{c^\prime c} \nonumber \\
            &= \frac{1}{2} \biggl\{ \text{Tr} (\boldsymbol{\Sigma}_c^{-1} \boldsymbol{\Sigma}_{c^\prime} + \boldsymbol{\Sigma}_{c^\prime}^{-1} \boldsymbol{\Sigma}_c) \nonumber \\
            &~~+ (\boldsymbol{\mu}_c - \boldsymbol{\mu}_{c^\prime})^t (\boldsymbol{\Sigma}_c^{-1} + \boldsymbol{\Sigma}_{c^\prime}^{-1}) (\boldsymbol{\mu}_c - \boldsymbol{\mu}_{c^\prime}) - 2d \biggr\}.
        \end{align}

@@

    Following OTB framework, two application are available. The first,
    called  \texttt{otbcli\_TrainGMMSelectionApp}, performs  a feature
    selection algorithm (SFS or SFFS) with the one given criterion (to
    choose  between  Jeffries-Matusita  dist.,  Kullback-Lieber  div.,
    Cohen’s  kappa, overall  accuracy,  mean  F1-score).  The  second,
    called \texttt{otbcli\_PredictGMMApp}, generates the thematic maps
    according to the  learn model. The only  technical limitation that
    the  training set  must  fit  in the  RAM  of  the computer.   The
    classification step is streamed and there is no limitation in term
    of image  size. The Figure~\ref{fig:otb:ffs} shows  a code excerpt
    to run the application.

>>>>

\IEEEPARstart{W}{ith} the increasing number of remote sensing missions, the quantity of available data for a given landscape becomes larger and larger. Satellite missions produce a huge amount of data on a regularly (daily) basis. From 2018, the EnMAP (Environmental Mapping and Analysis Program) satellites managed by the German space agency will produce images with 244 spectral bands with a resolution of 30x30m with a frequency of revisit of 4 days\cite{Müller09enmap}. The Hyperspectral Infrared Imager (HyspIRI) of NASA will also deliver images with 212 spectral bands. Additionally to hyperspectral data, the amount of available hypertemporal data increases a lot. For instance, the European satellites Sentinel-2 were launched successfully recently and 2 Terabytes of data will be released every days~\cite{drusch2012sentinel}. This high volume of earth observation data provides an accurate information of land state and functioning, and help to improve the understanding of the earth~\cite{rs1010001}. However, processing such data is more and more challenging because of statistical and computational issues.

In the spectral or temporal domain, a pixel is represented by a vector
for   which  each   component   corresponds   to  a   spectral/temporal
measurement.  The  size of  the vector  is the  number of  spectral or
temporal  measurements.   For  hyperspectral images,  this  number  is
typically   about   several   hundreds  while   for   the   Sentinel-2
multitemporal images, the number of spectro-temporal measurement for a
given  year  is approximately  one  thousand.   When working  in  high
dimensional  space,  statistical  methods  made for  low  or  moderate
dimensional  space do  not  adapt  well.  For  instance,  the rate  of
convergence of the statistical estimation decreases when the dimension
grows while conjointly the number of parameters to estimate increases,
making    the    estimation    of   the    model    parameters    very
difficult~\cite{donoho}.  Consequently,  with a limited  training set,
beyond a certain limit, the classification accuracy actually decreases
as the number of features increases~\cite{hughes}.  For the purpose of
classification,  these  problems are  related  to  the \emph{curse  of
  dimensionality}. This is a major drawback in remote sensing since it
is  difficult to  collect a  large and  an accurate  ground-truth.  An
intensive work has  been performed in the remote  sensing community to
build  accurate classifiers  for  high  dimensional images.   Bayesian
models~\cite{book:landgrebe}, feature extraction and feature reduction
techniques~\cite{book:landgrebe,DR:guided:tour},                random
forest~\cite{1396322},   neural  networks~\cite{5411821}   and  kernel
methods~\cite{kernel:methods:rs}  have   been  investigated   for  the
classification of such images.

The  size of  the data  is increasing  with respect  to the  number of
measurement per  pixel. The data  volume of an hyperspectral  image is
typically    several    Gigabytes     per    acquisition    ($\approx$
300km$^2$). Multitemporal data are  now available freely from internet
stream    (see    for    instance     the    Copernicus    data    hub
\url{https://cophub.copernicus.eu/}).  This very  large volume of data
requires   specific   computing  infrastructure.    High   performance
computing   is   actually   investigated   by   the   remote   sensing
community~\cite{christophe2011remote,plaza2011high}.  Main  issues are
related to the use of  parallel approach (multi-core, CPU, clusters) to
improve the  processing time, and  to the use of  streaming techniques
when data  does not fit in  memory.  One popular open  source software
solution  is  the Orfeo  Toolbox,  developed  by the  French  National
Agency~\cite{christophe2008orfeo}.   Streaming and  parallel computing
are  conveniently  proposed  to  the  user  through  several  modules
``\emph{ready to use}''.

@@

There are two ways to reduce dimension~\cite[Chapter 1]{Guyon:2006:FEF:1208773}:
features extraction and feature selection. Feature extraction means reformulate
and summarize the information by creating new features in combining the existing
ones, it is sometimes referred to \emph{feature construction}. Linear
combination of the initial features can be extracted using Principal Component
Analysis (PCA)~\cite{jimenez1998supervised} or Independent Component Analysis
\cite{villa2011hyperspectral}. Supervised extraction method has also been
investigated such as Fischer discriminant analysis and decision boundary feature
extraction~\cite{book:landgrebe}. To the contrary, features selection selects a
subset of existing features identified as the most relevant for a given
criterion. It has the additional advantage to be much more easily understandable for the end-user.

@@

    For mixture models, it is assumed  that a given sample $\mathbf{x}$ is
    the realization  of a  random vector which  distribution is  a mixture
    (convex     combination)     of      several     class     conditioned
    distribution~\cite{Fraley00model-basedclustering}:
    \begin{equation}
        p(\mathbf{x}) = \sum_{c=1}^{C} \pi_c f_c(\mathbf{x}|\theta)
    \end{equation}
    where $\pi_c$ is  the prior, \emph{i.e.}, the  proportion of class
    $c$  and  $f_c$  a  parametric density  function  control by
    $\theta$.

@@

        The    KL   divergence    is   not    symmetric,   \emph{i.e.}
        $KL_{cc^\prime} \ne KL_{c^\prime c}$.  A symmetrical version  is used to
        compute the criterion function:
        \begin{align}\label{eq:skl}
          SKL_{cc^\prime} &=KL_{cc^\prime} + KL_{c^\prime c} \nonumber \\
            &= \frac{1}{2} \biggl\{ \text{Tr} (\boldsymbol{\Sigma}_c^{-1} \boldsymbol{\Sigma}_{c^\prime} + \boldsymbol{\Sigma}_{c^\prime}^{-1} \boldsymbol{\Sigma}_c) \nonumber \\
            &~~+ (\boldsymbol{\mu}_c - \boldsymbol{\mu}_{c^\prime})^t (\boldsymbol{\Sigma}_c^{-1} + \boldsymbol{\Sigma}_{c^\prime}^{-1}) (\boldsymbol{\mu}_c - \boldsymbol{\mu}_{c^\prime}) - 2d \biggr\}.
        \end{align}

@@

    Following OTB framework, two application are available. The first,
    called  \texttt{otbcli\_TrainGMMSelectionApp}, performs  a feature
    selection algorithm (SFS or SFFS) with the one given criterion (to
    choose  between  Jeffries-Matusita  dist.,  Kullback-Leibler  div.,
    Cohen’s  kappa, overall  accuracy,  mean  F1-score).  The  second,
    called \texttt{otbcli\_PredictGMMApp}, generates the thematic maps
    according to the  learn model. The only  technical limitation that
    the  training set  must  fit  in the  RAM  of  the computer.   The
    classification step is streamed and there is no limitation in term
    of image  size. The Figure~\ref{fig:otb:ffs} shows  a code excerpt
    to run the application.
